\section{Efficiency Analysis}

Here we try to analyze the speedup of Parareal analytically, and then through
scalability studies conducted on the supercomputer Prince, here at NYU. 

\subsection{Theoretical Results}

We try to directly estimate the speedup of the Naive OpenMP implementation. We
denote our speedup $S$ by: (\cite{fieldstalk} slide 17)
\begin{equation*}
  S = \frac{\text{Time taken by fine solver}}{\text{Parareal time}}
\end{equation*}
Let the time taken by the fine and course solvers respectively be denoted by
$T_f$ and $T_g$. Furthermore, suppose we have the optimal amount of processors
$P$, such that $\Delta t = T/P$, where $T$ is the time to integrate up to.

Considering this, at each step parareal makes a predictor and corrector
computation. The predictor is just a run of the course solver, costing $PT_g$.
The corrector has two parts, the fine and course portion. The fine portion is
parallelized optimally, and therefore costs just $T_f$, instead of $PT_f$.
However, the course portion of the corrector takes advantage of the
\textit{first same as last} property, and doesn't have to be counted sepreately.
Recall, also, before the parareal iteration, we make an initial guess using one
course solve, $PT_g$. Therefore, for a parareal with $k$ iterations it would
have speedup:
\[
  S = \frac{PT_f}{PT_g + k(PT_g + T_f)} 
  = \frac{1}{\frac{T_g}{T_f} + k(\frac{T_g}{T_f} + \frac{1}{P})} 
  = \frac{1}{\frac{T_g}{T_f}(1+k) + \frac{k}{P}} 
\]
Notice, under the limit:
\[
  \lim_{P \to \infty} S = \frac{1}{\frac{T_g}{T_f}(1+k)} = \frac{T_f}{T_g(1+k)}
\]
Supposing we have the idealized case of $k = 1$, then we would find that this
would result in speedup $\frac{T_f}{2T_g}$, which, given that $T_g << T_f$,
could be significant. However, in practice, $k$ has to be taken to be large
enough so that $S$ isn't too great. This is one of the first indicators we have
of why $T_g << T_f$ must be true. In general, knowing that $T_g \leq T_f
\implies S \leq \frac{1}{k+1}$, which is somewhat dismal. 

\subsection{Scalability}
